The Nature of Breath: A Biomimetic Kinetic Sculpture That Breathes

### 2021-12-01 finalproject/report/ms1report.tex
new audio files and paper edits

\begin{abstract}
the nature of breath is...
\end{abstract}

\keywords{Keywords}

%------------------------------
\section{Introduction}

@@

Using the human act of breathing as a lens, this work attempts to lay a foundation for inquiry into the relationship between what is considered to be natural and what is or has been alive. While the focus of this work is on a specifically human form of breath, respiration in broader context is omnipresent in living organisms and is therefore an apt biological process with which to address this question.

@@

With the exception of cellular respiration, respiration in breathing organisms is a visibly mechanical process. Humans have created many mechanisms to mimic various aspects of the respiratory system for purposes ranging from medical devices, underwater exploration, to speech synthesis and more. Most of these devices are higly specialized and seek to mimic partifulatr aspects of respiration such as oxygen delivery, or vocal modulation. These works rarely mimic the actual mechanics of the components of the human body responsible for resporation favoring more abstracted and efficient mechanical systems to achieve their goal. Using the freedom afforded by kinetic sculpture, this work attempts to blend a human-biomimetic approach -- highlighting the mehcnical simplicity of the human body -- with a highly technical approach which allows for precise and complex control over the resulting breath.

This work also asks us to consider the fragility of human breath in light of the COVID-19 pandemic. Seemingly overnight, much of the world became obsessed with breath as the tool of the virus spread. In the two years since, we have become incredibly attuned to the quality and sound of our own breath and what it may say about our mortality. This work seeks to probe these newfound sensitivities, forcing us to question which actions may lead breath to be easy or sharpy painful.

@@

\subsection{temporal relevance}
% the pandemic...the soundscape of the pandemic...
Teh COVID-19 pandemic has caused a sort of crystalization on many of the aforementioned optics, and has

During the pandemic many people tried ot mkae ventilators at home. Segue...

\subsection{prior work and relevance}
% talk about the history of breathing apparatus and note that none mimic the human body
Mechanical ventilation has been around for a long time... Since it  seeks to provide "replacement of the respiratory muscles". \cite{ventilatorhistory} This results in apparati that resembly the respiratory system only in the sens that they cause a regutaled chnage in pressure, but have no need for recreating htep rocess by which breaht sounds are actually created. Simialrly, the cield of biomemtic robotics has little use for such inquiry - respiration is not necessary for an electronic humanoid and it is much easier to produce vocal noises in a digital manner.

Two notable works in the field of kinetic sculpture provide reference, Maywa Denkye's lauching machine and another guy's breath machine thing. These two pieces provoke thought about breath, but still use simple bellow bases mechanisms to produce airflow. %\cite{deynke} \cite{otherguy}

Since the focus of this piece is the noise of breath

\section{implementation}

\subsection{system breakdown}

While the avian respiratory system is, in fact, more complex, the human respiratory system is a very complicated biological system. \cite{breathcomplexity} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems. Each sub-system exists solely for the purpose of generating breath noises and not for the actual exchnage of oxygen as would normally occur during respiration.

@@

Actuation is the system which causes air to move through the respiratory system. In our own bodies this is acieved by the movemne of the diaphgragm. In ventilators this is ofent acieved by a pump or bellows.

@@

Modulation is the way the sound of breath is shaped as it leaves the body. This is acheived by the bronchial tree, trachea, larynx, and mouth. After actuation has been achieved, simulation of this system ist he most relevant and interesting part of this.

\subsection{making}

So far a few different protptypes ahve been attempted. Focusing on the two main areas of inquiry, there have been auditory prototypes -- exploring the effect os disembodied breath noises -- as well as machnical prototypes -- exploring ways to achieve a physiologically accurate breahting mechanism.

\subsubsection{auditory prototypes}

@@

\subsubsection{physical prototypes}

@@

\subsubsection{final prototypes}

@@

\section{further work}

there's always more to do...

@@

\section{conclusion}

The complexity of the human respiratory system lies not in its gross mechanics, but in the materials and molecular mechanics that allow for the transfer of oxyneg to the blood. The goal of this work was to isolate the mechanical components which form the noise created by breathing.

talk about the big distinction of artificial and natural

why it's important to think about this in the context of humanity's relationship with the natural world

@@

\section{Bibliography}
%The difference between a reference list and a bibliography is that in your references, you list all the sources you directly referred to in the body of your writing in numerical order, whereas a bibliography includes an alphabetical listing of all those authors and sources that you have consulted while writing your essay. Use the same format as for the references otherwise. Those using Latex will follow the usual cite command format.

\section{Author Biography}
Lauria Clarke is an engineer and artist living in New York. She completed an MS in Computer Engineering at Northeastern University in 2017 and is currently persuing an MFA in Design and Technology at The Parsons School of Design. She specializes in digital hardware design and kinetic sculpture and is always curious to learn more about humanity's relationship with the natural world.

@@

>>>>

\begin{abstract}
The Nature of Breath is a kinetic sculpture exploring the sound and mechanics of human breath. By mimicking the mechanics of the human body this piece asks what it means for a machine to breath, drawing an uncertain line between what is considered natural and artificial. The Nature of Breath is also situated in response to the COVID-19 pandemic. In response to the virus' spread over the last two years, humans around the globe have become unusually attuned to the degrees of health or sickness that breath can indicate as well as the factors that may change one to the other. 
\end{abstract}

\keywords{Keywords}
kinetic sculpture, COVID-19, biomimicry, breath
%------------------------------
\section{Introduction}

@@

Using the human act of breathing as a lens, this work attempts to lay a foundation for inquiry into the relationship between what is considered to be natural and what is or has been alive. While the focus of this work is on a specifically human form of breath, respiration in broader context is omnipresent in living organisms and is therefore an apt biological process with which to address this question. This machine may breath in a natural way, but its appearance clearly identifies it as something which is purely artificial.

@@

With the exception of cellular respiration, respiration in breathing organisms is a visibly mechanical process. Humans have created many mechanisms to mimic various aspects of the respiratory system for purposes ranging from medical devices, underwater exploration, to speech synthesis and more. Most of these devices are highly specialized and seek to mimic particular aspects of respiration such as oxygen delivery, or vocal modulation. These works rarely mimic the actual mechanics of the components of the human body responsible for respiration favoring more abstracted and efficient mechanical systems to achieve their goal. Using the freedom afforded by kinetic sculpture, this work attempts to blend a human-biomimetic approach -- highlighting the mechanical simplicity of the human body -- with a highly technical approach which allows for precise and complex control over the resulting breath.

This work also asks us to consider the fragility of human breath in light of the COVID-19 pandemic. Seemingly overnight, much of the world became obsessed with breath as the toll of the virus spread. In the two years since, we have become incredibly attuned to the quality and sound of our own breath and what it may say about our mortality. This work seeks to probe these newfound sensitivities, forcing us to question which actions may lead breath to be easy or sharply painful.

@@

\subsection{Prior Work and Relevance}
% talk about the history of breathing apparatus and note that none mimic the human body
Mechanical ventilation has been around for a long time (include specific reference)...it  seeks to provide "replacement of the respiratory muscles". \cite{ventilatorhistory} This results in apparatuses that resemble the respiratory system only in the sense that they cause a regulated change in pressure which allows for oxygenation of the blood. The devices have no need, however, for recreating the process by which breath sounds are actually created. The field of biomimetic robotics has little use for such inquiry either -- respiration is not necessary for an electronic being and it is much easier to produce vocal noises digitally than mechanically.

Two notable works in the field of kinetic sculpture provide some reference and context for this work, Maywa Denki's WAHHA GO GO and Rafael Lozano-Hemmer's Vicious Circular Breathing. \cite{denki} \cite{lozanohemmer} These two pieces provoke thought about breath, but still use simple bellow bases mechanisms to produce airflow.

% context of the pandemic ventilators

\section{Implementation}

\subsection{System Breakdown}

While the avian respiratory system is, in fact, more complex, the human respiratory system is a very complicated biological system. \cite{gasexchangers} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems. Each sub-system exists solely for the purpose of generating breath noises and not for the actual exchange of oxygen as would normally occur during respiration.

@@

Actuation is the system which causes air to move through the respiratory system. In our own bodies this is achieved by the movement of the diaphragm. In ventilators this is often achieved by a pump or bellows.

@@

Modulation is the way the sound of breath is shaped as it leaves the body. This is achieved by the bronchial tree, trachea, larynx, and mouth. After actuation has been achieved, simulation of this system is the most relevant and interesting part of this.

\subsection{Making and Testing}

So far a few different prototypes have been attempted. Focusing on the two main areas of inquiry, there have been auditory prototypes -- exploring the effect of disembodied breath noises -- as well as mechanical prototypes -- exploring ways to achieve a physiologically accurate breathing mechanism. In addition to creating a visual tie between the biological nature of the respiratory system, I chose to closely mimic the diaphragm and trachea in this way to prevent the noise of the mechanical components from overwhelming the noise of the breath produced. These approaches rely on material properties of -- elasticity and rigidity -- of latex and vinyl to be functional.

\subsubsection{Auditory Prototypes}

@@

\subsubsection{Physical Prototypes}

@@

\subsubsection{Final Prototypes}

@@

\section{Further Work}

well, there's always more to do...

@@

\section{Conclusion}

The complexity of the human respiratory system lies not in its gross mechanics, but in the materials and molecular mechanics that allow for the transfer of oxygen to the blood. The goal of this work was to isolate the mechanical components which form the noise created by breathing.

bring this back to the question of the relationship between what is natural and what is alive

here is where I can cite some of the exciting reading I've done and talk about the big distinction of artificial and natural

\cite{zimmer}
\cite{haraway}
\cite{smith}

% why it's important to think about this in the context of humanity's relationship with the natural world

@@

\section{Author Biography}
Lauria Clarke is an engineer and artist living in New York. She completed a MS in Computer Engineering at Northeastern University in 2017 and is currently pursuing an MFA in Design and Technology at The Parsons School of Design. She specializes in digital hardware design and kinetic sculpture and is always curious to learn more about humanity's relationship with the natural world.

@@

\section{Note to Harpreet}

Hi Harpreet,

My apologies, this is quite rough. I plan to finish fleshing out the prior work and relevance section and then collect and polish all the writing I've done about my prototypes for the implementation section. I accidentally used up a lot of time writing about the uncanny valley and it's relevance in biomimetics/biodesign/biotechnology...which was far outside the scope of this paper

Lauria
